---
title: "SARS-CoV-2: Visualization of long-read junctions."
---

# Load libraries
```{r}
library(tidyverse)
library(scales) #allows use of "comma" to replace i.e. 1e6 with 1,000,000 during plotting

library(pheatmap)
library(patchwork) # allows composing/orienting plots
library(RColorBrewer)

# geneplots
library(gggenes)


library(reshape2) #for melt function
library(ggrepel) # labeling points on a dotplot

library(hashmap) # use column as key for another column

library(ggrepel) # repelled labels


```

# gggenes tibbles
```{r}
sars2_orf1a_gene <- tibble(
  molecule = "SARS-CoV2_ORF1a",
  gene = c("ORF1a", "ORF1b", "S"),
  start = c(266, 13468, 21563),
  end = c(13483, 21555, 25384),
  strand = "forward",
  direction = 1
)

sars2_genes <- tibble(
  molecule = "SARS-CoV2",
  gene = c("S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"),
  start = c(21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274, 29558),
  end = c(25384, 26220, 26472, 27191, 27387, 27759, 27887, 28259, 29533, 29674),
  strand = "forward",
  direction = 1
)
# Specify order of gene for plotting
sars2_genes$gene <- factor(sars2_genes$gene, levels=c("S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"), ordered = T)

sars2_ALL_genes <- tibble(
  molecule = "SARS-CoV2",
  gene = c("ORF1a", "ORF1b","S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"),
  start = c(266, 13468, 21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274, 29558),
  end = c(13483, 21555,25384, 26220, 26472, 27191, 27387, 27759, 27887, 28259, 29533, 29674),
  strand = "forward",
  direction = 1
)
# Specify order of gene for plotting
sars2_ALL_genes$gene <- factor(sars2_ALL_genes$gene, levels=c("ORF1a", "ORF1b", "S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"), ordered = T)
```

# IMPORT dRNAseq (minimap2 mapped)
```{r}
taiaroa <- read_delim("inputs/taiaroa_mapped_junctions.tsv", delim="\t")
kim <- read_delim("inputs/kim_mapped_junctions.tsv", delim="\t")
davidson <- read_delim("inputs/davidson_mapped_junctions.tsv", delim="\t")

colnames(taiaroa) <- c("read_name", "five", "three")
colnames(kim) <- c("read_name", "five", "three")
colnames(davidson) <- c("read_name", "five", "three")


label_genes_by_3prime <- function(in_tbl) {
  in_tbl %>%
    mutate(gene = "Intergenic") %>%
    mutate(gene = ifelse(three>=21563, ifelse(three<=25384,"1. S", "Intergenic" ), "Intergenic")) %>%
    mutate(gene = ifelse(three>=25393, ifelse(three<=26220,"2. ORF3a", gene ), gene)) %>%
    mutate(gene = ifelse(three>=26245, ifelse(three<=26472,"3. E", gene ), gene)) %>%
    mutate(gene = ifelse(three>=26523, ifelse(three<=27191,"4. M", gene ), gene)) %>%
    mutate(gene = ifelse(three>=27202, ifelse(three<=27387,"5. ORF6", gene ), gene)) %>%
    mutate(gene = ifelse(three>=27394, ifelse(three<=27887,"6. ORF7a/b", gene ), gene)) %>%
    mutate(gene = ifelse(three>=27894, ifelse(three<=28259,"7. ORF8", gene ), gene)) %>%
    mutate(gene = ifelse(three>=28274, ifelse(three<=29533,"8. N", gene ), gene)) %>%
    mutate(gene = ifelse(three>=29558, ifelse(three<=29674,"9. ORF10", gene ), gene))
}



# Label genes by their 3' junctions
davidson <- label_genes_by_3prime(davidson)

taiaroa <- label_genes_by_3prime(taiaroa) 

kim <- label_genes_by_3prime(kim)


# Calculate junction length for each
davidson <- davidson %>%
  mutate(junction_length=three-five)

taiaroa <- taiaroa %>%
  mutate(junction_length=three-five)

kim <- kim %>%
  mutate(junction_length=three-five)
```

# Import and normalize Rouskin lab structures
```{r}

exposure_1 <- read_delim("inputs/rouskin_structure/GSM4656059_sars2vivo1_SARS2MN985325WA_1_29882_popavg_reacts.txt", delim="\t")


# Normalize
#----------------#

normalize_structure_values <- function(in_tbl) {
  
  # Find median of top 10% of mismatches
  top <- top_n(in_tbl, nrow(in_tbl)/20, Mismatches)
  top_median <- top %>%
    pull(Mismatches) %>%
    median()
  in_tbl$Mismatches <- in_tbl$Mismatches/top_median
  
  # Reduce to 1 anything that is greater than 1
  in_tbl <- in_tbl %>%
    mutate(Mismatches = ifelse(Mismatches>1, 1, Mismatches))
  
  # Return result
  in_tbl %>%
    select(Position, Mismatches)
}


exposure_1norm <- normalize_structure_values(structure_1)
```


# Correlate junctions wtih DMS values - only noncanonical
```{r}
# Make key for identifying TRS locations
trs_locations <- tibble(
  gene = c("S", "ORF3a", "E", "M", "ORF6", "ORF7a", "ORF8", "N"),
  TRS = c(21555, 25384, 26236, 26472, 27040, 27387, 27887, 28259)
)
trs_map <- hashmap(trs_locations$gene, trs_locations$TRS)

classify_nc_c <- function(in_tbl) {
  in_tbl %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["S"]]-15 & three < trs_map[["S"]]+15,
    "Canonical",
    ""
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["ORF3a"]]-15 & three < trs_map[["ORF3a"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["E"]]-15 & three < trs_map[["E"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["M"]]-15 & three < trs_map[["M"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["ORF6"]]-15 & three < trs_map[["ORF6"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["ORF7a"]]-15 & three < trs_map[["ORF7a"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["ORF8"]]-15 & three < trs_map[["ORF8"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(
    five < 100 & three > trs_map[["N"]]-15 & three < trs_map[["N"]]+15,
    "Canonical",
    classification
    )) %>%
  mutate(classification = ifelse(classification == "", "Non-canonical", classification)) %>%
  mutate(classification = ifelse(
    five < 49 | five > 89,
    "Non-canonical",
    classification
    )) # This last one makes sure 5' is within 20bp of TRS-L
}

kim_classified <- classify_nc_c(kim)
kim_noncanonical <- kim_classified %>%
  filter(classification == "Non-canonical")


kim_3tally <- kim_noncanonical %>%
  group_by(three) %>%
  tally() %>%
  arrange(-n) %>%
  mutate(percentage = n*100/sum(n))
colnames(kim_3tally) <- c("Position", "n", "percentage")


kim_5tally <- kim_noncanonical %>%
  group_by(five) %>%
  tally() %>%
  arrange(-n) %>%
  mutate(percentage = n*100/sum(n))
colnames(kim_5tally) <- c("Position", "n", "percentage")



kim_exposure_3 <- inner_join(kim_3tally, exposure_1norm, by="Position")
kim_exposure_5 <- inner_join(kim_5tally, exposure_1norm, by="Position")



cor(kim_exposure_3$n, kim_exposure_3$Mismatches, method = "spearman")
cor(kim_exposure_5$n, kim_exposure_5$Mismatches, method = "spearman")


kim_exposure_3_plt <- ggplot(kim_exposure_3) +
  geom_point(aes(x=n, y=Mismatches)) +
  scale_y_log10() + scale_x_log10() + annotation_logticks(sides="lb") + 
  ylab("DMS Signal") + xlab("Number of Non-Canonical\n3' Junctions") +
  theme_minimal(base_size=20)
kim_exposure_3_plt

kim_exposure_5plt <- ggplot(kim_exposure_5) +
  geom_point(aes(x=n, y=Mismatches)) +
  scale_y_log10() + scale_x_log10() + annotation_logticks(sides="lb") + 
  ylab("DMS Signal") + xlab("Number of Non-Canonical\n5' Junctions") +
  theme_minimal(base_size=20)
kim_exposure_5plt

kim_exposure_agg_plt <- kim_exposure_5plt + 
  kim_exposure_3_plt + theme(axis.text.y=element_blank(),
                             axis.ticks.y=element_blank(),
                             axis.title.y=element_blank())

ggsave("figures/structure/kim_exposure_agg_plt.pdf", kim_exposure_agg_plt, height = 10, width = 20, units="cm")
saveRDS(kim_exposure_agg_plt, "rds/structure/kim_exposure_agg_plt.rds")

### Above, should probably just do percentage of non-canonical junctions, rather than number.

### Really need to look at the folded, "structured/paired vs unstructured/unpaired" data, and do that analysis. But above shows me that there aren't more junctions on exposed bases (note - there can be exposed bases even in very structured regions.. so, more important to look at the structured vs unstructured)

```

# Import and process the dot structure
# dots = unpaired, parentheses = paired
```{r}
dot_structure <- read_delim("inputs/rouskin_structure/parsed_structures.txt", delim="\t")

dot_structure <- dot_structure %>%
  mutate(paired_status = ifelse(dot == ".", "Unpaired", "Paired"))
colnames(dot_structure)[1] <- c("Position")

### need to calculate structured vs unstructured stretches... that'll be more accurate. Requires python, because I'm not as good in R.

```


# Check association between dot structure and non-canonical junctions
```{r}
kim_5_dot <- dot_structure %>%
  inner_join(kim_5tally, by="Position")

kim_3_dot <- dot_structure %>%
  inner_join(kim_3tally, by="Position")


kim_5_dot %>% filter(structure_status != "mixed") %>%
  ggplot() +
  geom_boxplot(aes(x=structure_status, y=percentage)) + scale_y_log10() +
  geom_jitter(aes(x=structure_status, y=percentage))


kim_3_dot %>% filter(structure_status != "mixed") %>%
  ggplot() +
  geom_boxplot(aes(x=structure_status, y=percentage)) + scale_y_log10() +
  geom_jitter(aes(x=structure_status, y=percentage))



kim_5_structured_precentages <- kim_5_dot %>%
  filter(structure_status=="structured") %>%
  pull(percentage)
kim_5_unstructured_precentages <- kim_5_dot %>%
  filter(structure_status=="unstructured") %>%
  pull(percentage)

kim_3_structured_precentages <- kim_3_dot %>%
  filter(structure_status=="structured") %>%
  pull(percentage)
kim_3_unstructured_precentages <- kim_3_dot %>%
  filter(structure_status=="unstructured") %>%
  pull(percentage)

mean(kim_5_structured_precentages)
mean(kim_5_unstructured_precentages)

mean(kim_3_structured_precentages)
mean(kim_3_unstructured_precentages)

sum(kim_3_structured_precentages)
sum(kim_3_unstructured_precentages)

t.test(kim_5_structured_precentages, kim_5_unstructured_precentages)
t.test(kim_3_structured_precentages, kim_3_unstructured_precentages)



table(dot_structure$structure_status)
```


# Plot sum of nc-junction percentages in structured vs unstructured regions
```{r}

# Dot - get counts of structured/unstructured bases


# Sum percentages of structured and non-structured
kim_3_sums <- kim_3_dot %>% filter(structure_status != "mixed") %>%
  group_by(structure_status) %>%
  mutate(s = sum(percentage)) %>%
  distinct(structure_status, s)



kim_3_sums %>%
  ggplot()


tail(kim_3_dot, 1000)
```



# 
```{r}
dot_structure %>%
  ggplot() + 
  geom_point(aes(x=Position, y=structure_status))

# Compare structured and unstructred counts from 10K-20K and 20K-end

dot_structure %>%
  filter(Position <20000, Position >10000) %>%
  group_by(structure_status) %>% tally() %>%
  mutate(p = n*100/sum(n))
  
dot_structure %>%
  filter(Position >20000) %>%
  group_by(structure_status) %>% tally() %>%
  mutate(p = n*100/sum(n))




# Within N, 3x as much unstructured as structured regions
dot_structure %>%
  filter(Position >28274, Position < 29533) %>%
  group_by(structure_status) %>% tally() %>%
  mutate(p = n*100/sum(n))



# Within *N*, 10-fold as many junctions
# within unstructured regions as structured
kim_3_dot %>%
  filter(Position >28274, Position < 29533) %>%
  filter(structure_status=="unstructured") %>%
  pull(percentage) %>% sum
kim_3_dot %>%
  filter(Position >28274, Position < 29533) %>%
  filter(structure_status=="structured") %>%
  pull(percentage) %>% sum
```





# Lets look at 3' junctions...
```{r}

ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >= 21000),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status ),
               binwidth = 20) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Count") +
    geom_vline(xintercept = c(21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") +
    #scale_fill_manual(values=c("#BFC9CA", "#A93226", "#9B59B6", "#AED6F1", "#2471A3", "#48C9B0", "#1E8449", "#F7DC6F", "#AF601A" , "#5D6D7E")) +
    geom_gene_arrow(data = sars2_genes, aes(xmin = start, xmax = end, y = 5000, label=gene))


ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >28274, Position < 29533, structure_status!="mixed"),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") 
    #scale_fill_manual(values=c("#BFC9CA", "#A93226", "#9B59B6", "#AED6F1", "#2471A3", "#48C9B0", "#1E8449", "#F7DC6F", "#AF601A" , "#5D6D7E")) +
    #geom_gene_arrow(data = sars2_genes, aes(xmin = start, xmax = end, y = 5000, label=gene))


ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >28274, Position < 29533, structure_status=="mixed"),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") 

ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >28274, Position < 29533),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") 



kim_3_dot %>% filter(Position >28274, Position < 29533, structure_status!="mixed") %>% mutate(s = sum(n)) %>% pull(s) %>% unique()
kim_3_dot%>% filter(structure_status!="mixed") %>% mutate(s = sum(n)) %>% pull(s) %>% unique()
kim_3_dot%>% mutate(s = sum(n)) %>% pull(s) %>% unique()

```



# More 3' junctn by status
```{r}
# Full
ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >20000, Position < 29533),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") 

# only structured
ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >20000, Position < 29533, structure_status=="structured"),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") 


# only unstructured
# only structured
ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >20000, Position < 29533, structure_status=="unstructured"),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed")

# No mixed
# only unstructured
# only structured
ggplot() + 
geom_col(data=kim_3_dot %>% filter(Position >20000, Position < 29533, structure_status!="mixed"),
               aes(x=Position, y = percentage, color=structure_status, fill =structure_status )) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,1)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Percentage of non-canonical Junctions") +
    geom_vline(xintercept = c(28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed")

```


# TMP - what is the ORF6 TRS-mediated junction point?
```{r}
kim %>% filter(three > 26523, three < 27191) %>% group_by(three) %>% tally() %>%
  ggplot() + 
  geom_col(aes(x=three, y=n), color="black")

kim %>% filter(three > 26523, three < 27191) %>% group_by(three) %>% tally() %>% arrange(-n)

#ORF6 TRS location: 27040
```






# Correlate junctions wtih DMS values - including canonical
```{r}

# kim_3tally <- kim %>%
#   group_by(three) %>%
#   tally() %>%
#   arrange(-n)
# colnames(kim_3tally) <- c("Position", "n")
# 
# 
# kim_5tally <- kim %>%
#   group_by(five) %>%
#   tally() %>%
#   arrange(-n)
# colnames(kim_5tally) <- c("Position", "n")
# 
# 
# 
# kim_structure_3 <- inner_join(kim_3tally, structure_1norm, by="Position")
# kim_structure_5 <- inner_join(kim_5tally, structure_1norm, by="Position")
# 
# 
# 
# cor(kim_structure_3$n, kim_structure_3$Mismatches, method = "spearman")
# cor(kim_structure_5$n, kim_structure_5$Mismatches, method = "spearman")
# 
# 
# ggplot(kim_structure_3) +
#   geom_point(aes(x=n, y=Mismatches))
# 
# ggplot(kim_structure_5) +
#   geom_point(aes(x=n, y=Mismatches))
```