---
title: "SARS-CoV-2: Visualization of long-read junctions."
output: html_notebook
---

# Load libraries
```{r}
library(tidyverse)
library(scales) #allows use of "comma" to replace i.e. 1e6 with 1,000,000 during plotting

library(pheatmap)
library(patchwork) # allows composing/orienting plots
library(RColorBrewer)

# geneplots
library(gggenes)


library(reshape2) #for melt function
library(ggrepel) # labeling points on a dotplot

library(hashmap) # use column as key for another column

```

# gggenes tibbles
```{r}
sars2_orf1a_gene <- tibble(
  molecule = "SARS-CoV2_ORF1a",
  gene = c("ORF1a", "ORF1b", "S"),
  start = c(266, 13468, 21563),
  end = c(13483, 21555, 25384),
  strand = "forward",
  direction = 1
)

sars2_genes <- tibble(
  molecule = "SARS-CoV2",
  gene = c("S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"),
  start = c(21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274, 29558),
  end = c(25384, 26220, 26472, 27191, 27387, 27759, 27887, 28259, 29533, 29674),
  strand = "forward",
  direction = 1
)
# Specify order of gene for plotting
sars2_genes$gene <- factor(sars2_genes$gene, levels=c("S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"), ordered = T)

sars2_ALL_genes <- tibble(
  molecule = "SARS-CoV2",
  gene = c("ORF1a", "ORF1b","S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"),
  start = c(266, 13468, 21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274, 29558),
  end = c(13483, 21555,25384, 26220, 26472, 27191, 27387, 27759, 27887, 28259, 29533, 29674),
  strand = "forward",
  direction = 1
)
# Specify order of gene for plotting
sars2_ALL_genes$gene <- factor(sars2_ALL_genes$gene, levels=c("ORF1a", "ORF1b", "S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"), ordered = T)
```

# IMPORT dRNAseq (minimap2 mapped)
```{r}
taiaroa <- read_delim("inputs/taiaroa_mapped_junctions.tsv", delim="\t")
kim <- read_delim("inputs/kim_mapped_junctions.tsv", delim="\t")
davidson <- read_delim("inputs/davidson_mapped_junctions.tsv", delim="\t")

colnames(taiaroa) <- c("read_name", "five", "three")
colnames(kim) <- c("read_name", "five", "three")
colnames(davidson) <- c("read_name", "five", "three")


label_genes_by_3prime <- function(in_tbl) {
  in_tbl %>%
    mutate(gene = "Intergenic") %>%
    mutate(gene = ifelse(three>=21563, ifelse(three<=25384,"1. S", "Intergenic" ), "Intergenic")) %>%
    mutate(gene = ifelse(three>=25393, ifelse(three<=26220,"2. ORF3a", gene ), gene)) %>%
    mutate(gene = ifelse(three>=26245, ifelse(three<=26472,"3. E", gene ), gene)) %>%
    mutate(gene = ifelse(three>=26523, ifelse(three<=27191,"4. M", gene ), gene)) %>%
    mutate(gene = ifelse(three>=27202, ifelse(three<=27387,"5. ORF6", gene ), gene)) %>%
    mutate(gene = ifelse(three>=27394, ifelse(three<=27887,"6. ORF7a/b", gene ), gene)) %>%
    mutate(gene = ifelse(three>=27894, ifelse(three<=28259,"7. ORF8", gene ), gene)) %>%
    mutate(gene = ifelse(three>=28274, ifelse(three<=29533,"8. N", gene ), gene)) %>%
    mutate(gene = ifelse(three>=29558, ifelse(three<=29674,"9. ORF10", gene ), gene))
}



# Label genes by their 3' junctions
davidson <- label_genes_by_3prime(davidson)

taiaroa <- label_genes_by_3prime(taiaroa) 

kim <- label_genes_by_3prime(kim)


# Calculate junction length for each
davidson <- davidson %>%
  mutate(junction_length=three-five)

taiaroa <- taiaroa %>%
  mutate(junction_length=three-five)

kim <- kim %>%
  mutate(junction_length=three-five)
```

# IMPORT Blanco-Melo short read data (minimap2 mapped)
```{r}
SRR11517741 <- read_delim("inputs/SRR11517741_junctions.txt", delim="\t")
SRR11517742 <- read_delim("inputs/SRR11517742_junctions.txt", delim="\t")
SRR11517743 <- read_delim("inputs/SRR11517743_junctions.txt", delim="\t")


colnames(SRR11517741) <- c("read_name", "five", "three")
colnames(SRR11517742) <- c("read_name", "five", "three")
colnames(SRR11517743) <- c("read_name", "five", "three")

SRR11517741 <- label_genes_by_3prime(SRR11517741)
SRR11517742 <- label_genes_by_3prime(SRR11517742)
SRR11517743 <- label_genes_by_3prime(SRR11517743)
```

# IMPORT Blanco-Melo ferret data (minimap2 mapped)
```{r}
#Importing
ferret_cov <- read_delim("inputs/blanco_melo_ferret/SRR11517855_8_ferret.cov", delim="\t", col_names=c("genome", "position", "coverage"))
ferret_junc <- read_delim("inputs/blanco_melo_ferret/SRR11517855_8_ferret.junctions.txt", delim="\t")
colnames(ferret_junc) <- c("name", 'five', 'three')
```


# IMPORT Finkel 4hpi, 24hpi (STAR mapped)
```{r}
# Import
finkel_5hpi_1_cov <- read_delim("inputs/finkel/mRNA5_1.cov", delim="\t", col_names=c("genome", "position", "coverage"))
finkel_5hpi_2_cov <- read_delim("inputs/finkel/mRNA5_2.cov", delim="\t", col_names=c("genome", "position", "coverage"))
finkel_24hpi_1_cov <- read_delim("inputs/finkel/mRNA24_1.cov", delim="\t", col_names=c("genome", "position", "coverage"))
finkel_24hpi_2_cov <- read_delim("inputs/finkel/mRNA24_2.cov", delim="\t", col_names=c("genome", "position", "coverage"))


finkel_5hpi_1_junc <- read_delim("inputs/finkel/mRNA5_1.junctions.txt", delim="\t")
finkel_5hpi_2_junc <- read_delim("inputs/finkel/mRNA5_2.junctions.txt", delim="\t")
finkel_24hpi_1_junc <- read_delim("inputs/finkel/mRNA24_1.junctions.txt", delim="\t")
finkel_24hpi_2_junc <- read_delim("inputs/finkel/mRNA24_2.junctions.txt", delim="\t")
colnames(finkel_5hpi_1_junc) <- c("name", 'five', 'three')
colnames(finkel_5hpi_2_junc) <- c("name", 'five', 'three')
colnames(finkel_24hpi_1_junc) <- c("name", 'five', 'three')
colnames(finkel_24hpi_2_junc) <- c("name", 'five', 'three')
```

# IMPORT Emanuel 4hpi, 24hpi (STAR mapped)
```{r}
emanual_total_4hpi_1_cov <- read_delim("inputs/emanual_totalRNA/SRR11550047.cov", delim="\t", col_names=c("genome", "position", "coverage"))
emanual_total_4hpi_2_cov <- read_delim("inputs/emanual_totalRNA/SRR11550048.cov", delim="\t", col_names=c("genome", "position", "coverage"))
emanual_total_12hpi_1_cov <- read_delim("inputs/emanual_totalRNA/SRR11550043.cov", delim="\t", col_names=c("genome", "position", "coverage"))
emanual_total_12hpi_2_cov <- read_delim("inputs/emanual_totalRNA/SRR11550044.cov", delim="\t", col_names=c("genome", "position", "coverage"))
emanual_total_24hpi_1_cov <- read_delim("inputs/emanual_totalRNA/SRR11550045.cov", delim="\t", col_names=c("genome", "position", "coverage"))
emanual_total_24hpi_2_cov <- read_delim("inputs/emanual_totalRNA/SRR11550046.cov", delim="\t", col_names=c("genome", "position", "coverage"))

emanual_total_4hpi_1_junc <- read_delim("inputs/emanual_totalRNA/SRR11550047.junctions.txt", delim="\t")
emanual_total_4hpi_2_junc <- read_delim("inputs/emanual_totalRNA/SRR11550048.junctions.txt", delim="\t")
emanual_total_12hpi_1_junc <- read_delim("inputs/emanual_totalRNA/SRR11550043.junctions.txt", delim="\t")
emanual_total_12hpi_2_junc <- read_delim("inputs/emanual_totalRNA/SRR11550044.junctions.txt", delim="\t")
emanual_total_24hpi_1_junc <- read_delim("inputs/emanual_totalRNA/SRR11550045.junctions.txt", delim="\t")
emanual_total_24hpi_2_junc <- read_delim("inputs/emanual_totalRNA/SRR11550046.junctions.txt", delim="\t")
colnames(emanual_total_4hpi_1_junc) <- c("name", 'five', 'three')
colnames(emanual_total_4hpi_2_junc) <- c("name", 'five', 'three')
colnames(emanual_total_12hpi_1_junc) <- c("name", 'five', 'three')
colnames(emanual_total_12hpi_2_junc) <- c("name", 'five', 'three')
colnames(emanual_total_24hpi_1_junc) <- c("name", 'five', 'three')
colnames(emanual_total_24hpi_2_junc) <- c("name", 'five', 'three')
```
# IMPORT Emmanual polyA timecourse (STAR mapped)
```{r}

```


# IMPORT Suzuki organoids (minimap2 mapped)
```{r}
# Import
suzuki_bronchial_5dpi_1_cov <- read_delim("inputs/suzuki_bronchial_organoids/minimap2/SRR11811022.cov", delim="\t", col_names=c("genome", "position", "coverage"))
suzuki_bronchial_5dpi_2_cov <- read_delim("inputs/suzuki_bronchial_organoids/minimap2/SRR11811023.cov", delim="\t", col_names=c("genome", "position", "coverage"))
suzuki_bronchial_5dpi_3_cov <- read_delim("inputs/suzuki_bronchial_organoids/minimap2/SRR11811024.cov", delim="\t", col_names=c("genome", "position", "coverage"))


suzuki_bronchial_5dpi_1_junc <- read_delim("inputs/suzuki_bronchial_organoids/minimap2/SRR11811022.junctions.txt", delim="\t")
suzuki_bronchial_5dpi_2_junc <- read_delim("inputs/suzuki_bronchial_organoids/minimap2/SRR11811023.junctions.txt", delim="\t")
suzuki_bronchial_5dpi_3_junc <- read_delim("inputs/suzuki_bronchial_organoids/minimap2/SRR11811024.junctions.txt", delim="\t")

colnames(suzuki_bronchial_5dpi_1_junc) <- c("name", 'five', 'three')
colnames(suzuki_bronchial_5dpi_2_junc) <- c("name", 'five', 'three')
colnames(suzuki_bronchial_5dpi_3_junc) <- c("name", 'five', 'three')
```


# 2D histogram of junctions
```{r}

# all junctions
#-----------------------#
plot_all_junctions_2d_histogram <- function(in_tbl, max=100000) {
  ggplot(in_tbl, aes(three, five)) +
  stat_bin2d(binwidth = c(100,100) ) + 
  scale_fill_gradientn(limits=c(1,max),
                       colours=brewer.pal(5,"OrRd"),
                       trans="log10",
                       labels=comma,
                       breaks=10**(1:10)) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank()) +
  ylab("5'") + 
  xlab("3'") 
}

taiaroa_all_junctions <- plot_all_junctions_2d_histogram(taiaroa)
taiaroa_all_junctions

kim_all_junctions <- plot_all_junctions_2d_histogram(kim)
kim_all_junctions

davidson_all_junctions <- plot_all_junctions_2d_histogram(davidson)
davidson_all_junctions




#ggsave("figures/davidson_all_junctions.pdf", davidson_all_junctions, width=10, height=10)
```

# Long read - 5' 100bp  2d histograms
```{r}

# Plotting function
plot_5_100bp_2d_histogram <- function(in_tbl, max=100000) {
  ggplot(in_tbl %>%filter(five<=100, three>21000), aes(three, five)) +
  stat_bin2d(binwidth = c(100,1) ) + 
  scale_fill_gradientn(limits=c(1,max),
                       colours=brewer.pal(5,"OrRd"),
                       trans="log10",
                       labels=comma,
                       breaks=10**(1:10)) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank()) +
  ylab("5'") + 
  xlab("3'") +
  geom_vline(xintercept = c(21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274,29558) , 
              color = "black",
              size=0.1,
              alpha=0.5,
              linetype="dashed")
}

taiaroa_inset <- plot_5_100bp_2d_histogram(taiaroa)
kim_inset <- plot_5_100bp_2d_histogram(kim)
davidson_inset <- plot_5_100bp_2d_histogram(davidson)
```


# Long read 3' junctions
```{r}


plot_cannonical_3prime_junctions <- function(in_tbl, ymax=110000) {
  ggplot() +
    geom_histogram(data=in_tbl%>%filter(three>=21000, five<=100), aes(x=three), binwidth = 20) +
    theme_minimal() + 
    scale_y_continuous(limits=c(0,ymax)) +
    theme(panel.grid.minor = element_blank()) +
    theme(plot.title = element_text(size=10),
          axis.text.x = element_text(size = 10, angle = 45, hjust = .5, vjust = .5)) +
    guides(fill=guide_legend(title="Gene")) +
    xlab("3'") + 
    ylab("Count") +
    geom_vline(xintercept = c(21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274,29558) , 
                color = "black",
                size=0.2,
                alpha=0.5,
                linetype="dashed") +
    scale_fill_manual(values=c("#BFC9CA", "#A93226", "#9B59B6", "#AED6F1", "#2471A3", "#48C9B0", "#1E8449", "#F7DC6F", "#AF601A" , "#5D6D7E")) +
    geom_gene_arrow(data = sars2_genes, aes(xmin = start, xmax = end, y = ymax, fill = gene, label=gene))
  
}

taiaroa_3junctions_cannonical <- plot_cannonical_3prime_junctions(taiaroa, 110000)
taiaroa_3junctions_cannonical


kim_3junctions_cannonical <-  plot_cannonical_3prime_junctions(kim, 110000)
kim_3junctions_cannonical

davidson_3junctions_cannonical <- plot_cannonical_3prime_junctions(davidson, 110000)
davidson_3junctions_cannonical


```


# Plotting the short read junctions
```{r}

# All junctions
#--------------------#
SRR11517741_all_junctions <- plot_all_junctions_2d_histogram(SRR11517741, 500000)
SRR11517741_all_junctions

SRR11517742_all_junctions <- plot_all_junctions_2d_histogram(SRR11517742, 500000)
SRR11517742_all_junctions

SRR11517743_all_junctions <- plot_all_junctions_2d_histogram(SRR11517743, 500000)
SRR11517743_all_junctions

# Adding inset
SRR11517741_inset <- plot_5_100bp_2d_histogram(SRR11517741, 500000)
SRR11517741_inset

SRR11517742_inset <- plot_5_100bp_2d_histogram(SRR11517742, 500000)
SRR11517742_inset

SRR11517743_inset <- plot_5_100bp_2d_histogram(SRR11517743, 500000)
SRR11517743_inset

# 3' junctions
#--------------------#
SRR11517741_3junctions_cannonical <- plot_cannonical_3prime_junctions(SRR11517741, 35000)
SRR11517741_3junctions_cannonical

SRR11517742_3junctions_cannonical <- plot_cannonical_3prime_junctions(SRR11517742, 35000)
SRR11517742_3junctions_cannonical

SRR11517743_3junctions_cannonical <- plot_cannonical_3prime_junctions(SRR11517743, 35000)
SRR11517743_3junctions_cannonical
```



# Comnbining plots. Will do long- and short- reads separately.
```{r}

# Long reads
#----------------#

# Formating the insets
inset1 <- taiaroa_inset
inset2 <- kim_inset + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank())
inset3 <- taiaroa_inset + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank())

# Combining the 5-by-3 junctions
all_junctions_combined <- 
  taiaroa_all_junctions + 
  kim_all_junctions + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank()) + 
  davidson_all_junctions + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank()) + 
  inset1 + inset2 + inset3 +  
  plot_layout(guides = "collect", ncol=3)

# Combining cannonical
cannonical1 <- taiaroa_3junctions_cannonical + theme(plot.title = element_blank())
cannonical2 <-  kim_3junctions_cannonical + theme(axis.title.y=element_blank(),
                              axis.text.y=element_blank(),
                              plot.title = element_blank())
cannonical3 <- davidson_3junctions_cannonical + theme(axis.title.y=element_blank(),
                              axis.text.y=element_blank(),
                              plot.title = element_blank())


canonical <- cannonical1 + cannonical2 + cannonical3 +
  plot_layout(guides = "collect", ncol=3)

# Final assembled figure!!
figure1_junctions <- 
  all_junctions_combined /canonical
ggsave("figure1_junctions.pdf", figure1_junctions, width = 20, height=20, units="cm")
ggsave("figure1_junctions.svg", figure1_junctions, width = 20, height=20, units="cm")

```


# Assembling tenOever illumina junction figures
```{r}


# Formating the insets
short_inset1 <- SRR11517741_inset
short_inset2 <- SRR11517742_inset + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank())
short_inset3 <- SRR11517743_inset + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank())

# Combining the 5-by-3 junctions
short_all_junctions_combined <- 
  SRR11517741_all_junctions + 
  SRR11517742_all_junctions + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank()) + 
  SRR11517743_all_junctions + theme(axis.title.y=element_blank(),
                            axis.text.y=element_blank()) + 
  short_inset1 + short_inset2 + short_inset3 +  
  plot_layout(guides = "collect", ncol=3)


# Combining cannonical
short_cannonical1 <- SRR11517741_3junctions_cannonical + theme(plot.title = element_blank())
short_cannonical2 <-  SRR11517742_3junctions_cannonical + theme(axis.title.y=element_blank(),
                              axis.text.y=element_blank(),
                              plot.title = element_blank())
short_cannonical3 <- SRR11517743_3junctions_cannonical + theme(axis.title.y=element_blank(),
                              axis.text.y=element_blank(),
                              plot.title = element_blank())
short_canonical <- short_cannonical1 + short_cannonical2 + short_cannonical3 +
  plot_layout(guides = "collect", ncol=3)


# Assemble everything
short_junctions <- 
  short_all_junctions_combined /short_canonical

ggsave("short_junctions.pdf", short_junctions, width = 20, height=20, units="cm")
ggsave("short_junctions.svg", short_junctions, width = 20, height=20, units="cm")
```


# Importing coverage
```{r}
taiaroa_cov <- read_delim("inputs/taiaroa_mapped_mapped.cov", delim="\t", col_names=c("genome", "position", "coverage"))
kim_cov <- read_delim("inputs/kim_mapped_mapped.cov", delim="\t", col_names=c("genome", "position", "coverage"))
davidson_cov <- read_delim("inputs/davidson_mapped_mapped.cov", delim="\t", col_names=c("genome", "position", "coverage"))

SRR11517741_cov <- read_delim("inputs/SRR11517741.cov", delim="\t", col_names=c("genome", "position", "coverage"))
SRR11517742_cov <- read_delim("inputs/SRR11517742.cov", delim="\t", col_names=c("genome", "position", "coverage"))
SRR11517743_cov <- read_delim("inputs/SRR11517743.cov", delim="\t", col_names=c("genome", "position", "coverage"))

 
```



# Option 1 - Plotting ORF1a junctions vs coverage for 3short read and 3 long read datasets
```{r}

# Make gene map including only ORF1a, 1b, and S
sars2_orf1a_gene <- tibble(
  molecule = "SARS-CoV2_ORF1a",
  gene = c("ORF1a", "ORF1b", "S"),
  start = c(266, 13468, 21563),
  end = c(13483, 21555, 25384),
  strand = "forward",
  direction = 1
)


plot_coverage_and_juncs <- function(juncs, cov, arrow_height=200000, ymax=200000){
  ggplot() +
    stat_bin(data=juncs%>%filter(five>=240, five<=25384),
                   aes(x=five, y=cumsum(..count..)),
                   geom='line', 
                   color="red") +
    theme_minimal() + ylab("Count") + xlab("5'") +
    geom_line(data=cov %>% filter(position<=25384),
              aes(x=position, y=coverage)) +
    geom_gene_arrow(data = sars2_orf1a_gene, aes(xmin = start, xmax = end, y = arrow_height, fill = gene, label=gene)) +
    scale_fill_manual(values=c("#7DCEA0", "#F6DDCC", "#BFC9CA")) +
    geom_vline(xintercept = c(226, 13468, 13483, 21555, 25384) , 
                color = "black",
                size=0.2,
                alpha=0.8,
                linetype="dashed") +
  scale_y_continuous(labels=comma) +
  coord_cartesian(ylim=c(1,ymax)) +
    theme(axis.title.x = element_blank())
}

taiaroa_ORF1a_cov_and_juncs <- plot_coverage_and_juncs(taiaroa, taiaroa_cov, 5000, 5350)
kim_ORF1a_cov_and_juncs <- plot_coverage_and_juncs(kim, kim_cov, 25000, 26000)
davidson_ORF1a_cov_and_juncs <- plot_coverage_and_juncs(davidson, davidson_cov, 20000, 20500)

SRR11517741_ORF1a_cov_and_juncs <- plot_coverage_and_juncs(SRR11517741, SRR11517741_cov, 280000, 284000)
SRR11517742_ORF1a_cov_and_juncs <- plot_coverage_and_juncs(SRR11517742, SRR11517742_cov, 65000, 66500)
SRR11517743_ORF1a_cov_and_juncs <- plot_coverage_and_juncs(SRR11517743, SRR11517743_cov, 60000, 61500)

```


# Because I can't predict ORFs, from short-read data, lets at least look at where the 1A-originating junctions end up.
```{r}
plot_ORF1a_originating_3prime_junctions <- function(in_tbl, arrow_heigh=30000, ymax=30000) {

ggplot() +
    geom_histogram(data=in_tbl%>%filter(five>=240, five<=13483, three>21000),
                   aes(x=three),
                   binwidth = 30) +
      scale_fill_manual(values=c("#BFC9CA", "#A93226", "#9B59B6", "#AED6F1", "#2471A3", "#48C9B0", "#1E8449", "#F7DC6F", "#AF601A" , "#5D6D7E"))  +
    theme_minimal() + ylab("Count") + xlab("3'") +
  geom_gene_arrow(data = sars2_genes, aes(xmin = start, xmax = end, y = arrow_heigh, fill = gene, label=gene)) +
  geom_vline(xintercept = c(21563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274,29558) , 
              color = "black",
              size=0.2,
              alpha=0.8,
              linetype="dashed") +
  coord_cartesian(ylim=c(1,ymax)) +
    theme(axis.title.x = element_blank())
  
}

SRR11517741_ORF1a_originating <- plot_ORF1a_originating_3prime_junctions(SRR11517741, 18000, 18500)
SRR11517742_ORF1a_originating <- plot_ORF1a_originating_3prime_junctions(SRR11517743, 5000, 5200)
SRR11517743_ORF1a_originating <- plot_ORF1a_originating_3prime_junctions(SRR11517743, 5000, 5200)

taiaroa_ORF1a_originating <- plot_ORF1a_originating_3prime_junctions(taiaroa, 250, 260)
kim_ORF1a_originating <- plot_ORF1a_originating_3prime_junctions(kim, 1000, 1075)
davidson_ORF1a_originating <- plot_ORF1a_originating_3prime_junctions(davidson, 1200, 1275)

```

# Assembling ORF1a junction/cov figures
```{r}
ORF1a_cov_and_juncs_arranged <- taiaroa_ORF1a_cov_and_juncs + taiaroa_ORF1a_originating + theme(axis.title.y=element_blank()) + 
  kim_ORF1a_cov_and_juncs + kim_ORF1a_originating + theme(axis.title.y=element_blank()) +
  davidson_ORF1a_cov_and_juncs + davidson_ORF1a_originating + theme(axis.title.y=element_blank()) + 
  SRR11517741_ORF1a_cov_and_juncs + SRR11517741_ORF1a_originating + theme(axis.title.y=element_blank()) +
  SRR11517742_ORF1a_cov_and_juncs + SRR11517742_ORF1a_originating + theme(axis.title.y=element_blank()) +
  SRR11517743_ORF1a_cov_and_juncs + SRR11517743_ORF1a_originating + theme(axis.title.y=element_blank()) +
  plot_layout(guides = "collect", ncol=2, nrow=6)

ggsave("ORF1a_cov_and_juncs.pdf", ORF1a_cov_and_juncs_arranged, height=25, width=25, units="cm")
ggsave("ORF1a_cov_and_juncs.svg", ORF1a_cov_and_juncs_arranged, height=25, width=25, units="cm")
```

# Zoom-in of short reads cov and juncs chart for supplement
```{r}
short_zoomed_cov_and_juncs_arranged <- SRR11517741_ORF1a_cov_and_juncs + coord_cartesian(ylim=c(1, 120000)) +

SRR11517742_ORF1a_cov_and_juncs + coord_cartesian(ylim=c(1, 25000)) + 

SRR11517743_ORF1a_cov_and_juncs + coord_cartesian(ylim=c(1, 25000)) & theme(legend.position="NA")

ggsave("ORF1a_cov_and_juncs_SHORT_ZOOMED.pdf", short_zoomed_cov_and_juncs_arranged, height=10, width=20, units="cm")
ggsave("ORF1a_cov_and_juncs_SHORT_ZOOMED.svg", short_zoomed_cov_and_juncs_arranged, height=10, width=20, units="cm")
```


# Importing junction and coverage for CoV 229E
```{r}
CoV_229E_juncs <- read_delim("inputs/229E_long_junctions.txt", delim="\t")
colnames(CoV_229E_juncs) <- c("read_name", "five", "three")

CoV_229E_cov <- read_delim("inputs/229E_long.cov", delim="\t", col_names=c("genome", "position", "coverage"))
```

# Plotting junctions and coverage for CoV 229E
```{r}

# Make gene map including only ORF1a, 1b, and S
CoV_229E_orf1a_gene <- tibble(
  molecule = "SARS-CoV2_ORF1a",
  gene = c("ORF1a", "ORF1b", "S"),
  start = c(293, 12520, 20570),
  end = c(12520, 20568, 24091),
  strand = "forward",
  direction = 1
)

plot_coverage_and_juncs_229E <- function(juncs, cov, arrow_height=200000, ymax=200000){
  ggplot() +
    stat_bin(data=juncs%>%filter(five>=293, five<=24091),
                   aes(x=five, y=cumsum(..count..)),
                   geom='line', 
                   color="red") +
    theme_minimal() + ylab("Count") + xlab("5'") +
    geom_line(data=cov %>% filter(position<=24091),
              aes(x=position, y=coverage)) +
    geom_gene_arrow(data = CoV_229E_orf1a_gene, aes(xmin = start, xmax = end, y = arrow_height, fill = gene, label=gene)) +
    scale_fill_manual(values=c("#7DCEA0", "#F6DDCC", "#BFC9CA")) +
    geom_vline(xintercept = c(293, 12520, 12520, 20568, 20570, 24091) , 
                color = "black",
                size=0.2,
                alpha=0.8,
                linetype="dashed") +
  scale_y_continuous(labels=comma) +
  coord_cartesian(ylim=c(1,ymax))
}

CoV_229E_cov_and_juncs_plt <- plot_coverage_and_juncs_229E(CoV_229E_juncs, CoV_229E_cov, 16000, 16000) 

#geom_vline(xintercept=c(620, 3200), color="Blue", linetype="dashed")
# First inflection happens just after nsp1 around position 620, and second inflection happens just after nsp2 around 3200

ggsave("229E_cov_and_juncs.pdf", CoV_229E_cov_and_juncs_plt, height=10, width=15, units="cm")
ggsave("229E_cov_and_juncs.svg", CoV_229E_cov_and_juncs_plt, height=10, width=15, units="cm")
```



# gggenes virus genome map and subgenomic RNAs
```{r}

 # NOTE - FOR THIS ONE HAVE SQUEEZED IN ORF1AB FOR USE PLOTTING THE SUBGENOMIC RNAS RELATIVE TO FULL GENOME
#-----------------------------------#

all_sars2_genes <- c("Leader", "ORF1AB", "S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10")
TMP_sars2_genome <- tibble(
  molecule = "SARS-CoV2",
  gene = factor(all_sars2_genes, levels=c("Leader", "ORF1AB", "S", "ORF3a", "E", "M", "ORF6", "ORF7A", "ORF7B", "ORF8", "N", "ORF10"), ordered = T),
  start = c(19735, 20000, 23563, 25393, 26245, 26523, 27202, 27394, 27756, 27894, 28274, 29558),
  end = c(19800, 23555, 25384, 26220, 26472, 27191, 27387, 27759, 27887, 28259, 29533, 29674),
  strand = "forward",
  direction = 1
)

map_colors <- c("#000000", "#F6DDCC", "#BFC9CA", "#A93226", "#9B59B6", "#AED6F1", "#2471A3", "#48C9B0", "#1E8449", "#F7DC6F", "#AF601A" , "#5D6D7E")

genome_map_plt <- ggplot(TMP_sars2_genome, aes(xmin = start, xmax = end, y =
                                            molecule, fill = gene, label = gene)) +
  geom_gene_arrow(arrowhead_height = unit(3, "mm"), arrowhead_width = unit(1, "mm")) +
  theme_genes() +
  scale_fill_manual(values=map_colors)
ggsave("genome_maps/genome_map_condensed.pdf", genome_map_plt)
ggsave("genome_maps/genome_map_condensed.svg", genome_map_plt)


# Make map with the non-ORF1ab genes, and label the TRS'
#-----------------------------------#
sars2_genes_plt_TRS_lines <- ggplot(sars2_genes, aes(xmin = start, xmax = end, y =
                                            molecule, fill = gene, label = gene)) +
  geom_gene_arrow(arrowhead_height = unit(3, "mm"), arrowhead_width = unit(1, "mm")) +
  theme_genes() +
  scale_fill_manual(values=map_colors[3:length(map_colors)]) + 
  geom_vline(xintercept=c(21555, 25384,26236, 26472, 27040, 27387, 27887, 28259))
ggsave("genome_maps/genes_w_TRS_lines.pdf", sars2_genes_plt_TRS_lines)
ggsave("genome_maps/genes_w_TRS_lines.svg", sars2_genes_plt_TRS_lines)

```



